---
title: "Assignment 3"
author: "--"
date: "October 24, 2017"
output: 
  prettydoc::html_pretty:
  theme: cayman
  highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1



```{r}

```

## Question 2

If X and Y are random variables and:

$\mu_{X} =  E(X) = 1$, $\mu_{Y} =  E(Y) = 2$

$\sigma_{x} =  SD(X) = 3$, $\sigma_{Y} =  SD(Y) = 4$

$\rho_{XY} = CORR(X,Y) = 0.5$

Since we know $\rho_{XY}$ and we know that $\rho_{XY} = \frac{COV(X,Y)}{\sigma_{X} \sigma_{Y}}$ we can find $COV(X,Y)$:

$COV(X,Y) = \sigma_{X} \sigma_{Y} \rho_{XY} = 6$

#### a) 

$E(2X - Y + 5)$
$= E(2X) + E(-Y) + E(5)$
$= 2E(X) - E(Y) + E(5)$
$= 2 - 2 + 5$
$= 5$

Therefore $E(2X - Y + 5) = 5$

#### b) 

$VAR(2X - Y + 5)$
$= VAR(2X) + VAR(-Y) + VAR(5) + 2COV(2X,-Y) + 2COV(2X,5) + 2COV(-Y,5)$
$= VAR(2X) + VAR(-Y) + 2COV(2X,-Y)$
$= E((2X - 2\mu_{X})^2) + E(((-1)y - (-1)\mu_{Y})^2) + 2E((2X - 2\mu_{X})((-1)Y - (-1)\mu_{Y}))$
$= E(4(X-\mu_{X})^2) + E((1)(Y-\mu_{Y})^2) + 2E(2(X-\mu_{X})(-1)(Y-\mu_{Y}))$
$= 4E((X-\mu_{X})^2) + E((Y - \mu_{Y})^2) - 4E((X-\mu_{X})(Y-\mu_{Y}))$
$= 4\sigma^2_{X} + \sigma^2_{Y} - 4COV(X,Y)$
$= 4(9) + 16 - 4(6)$
$= 28$

$SD(2X - Y + 5) = \sqrt{VAR(2X - Y + 5)}$
$SD(2X - Y + 5) = 5.291503 \approx 5.29$



## Question 3

$\mu_{x} = \frac{1}{3}$, $\sigma^2_X=\frac{1}{18}$

We are assuming that $X_{1}, X_{2}, ...,X_{n}$ are iie (independent and identically distributed).

If $T = X_{1} + X{2} + X{3} + ... + X_{90} = \sum\limits^{90}_{i=1} X_{i}$ Then:

$E(T) = E(\sum\limits^{90}_{i=1} X_{i}) = \sum\limits^{90}_{i=1} X = nX = 60$

$VAR(T) = \sum\limits^{90}_{i=1} VAR(X) = \frac{90}{18} = 5$

By central limit theorem, T will be distributed $~N(60, \sqrt{5})$

Now, if we would like to find $P(T>65)$ we can use the standard normal distribtuion to easily find an approximate probability.

$Z= \frac{\bar{x} - \mu_{T}}{\sigma_{T}} = \frac{65-60}{\sqrt{5}} = 2.236$

The corresponding probability for $Z=2.236$ is $(1-\alpha) = 0.01257$. Therefore the probability that T is greater than 65, $P(T>65) \approx 0.01257$.

## Question 4

#### a) 

#### b)

The plot for the above distributution is as follows:
```{r}

```