---
title: "Question 2"
author: "Addison van den Hoeven, Mona Philippon, Deepwinder Singh"
date: "October 18, 2017"
output:
  prettydoc::html_pretty:
  theme: cayman
  highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### a) Calculate probability that a normal random variable with mean 22 and variance 25

$\sigma = \sqrt{25} = 5$
$\mu = 22$

```{r}
sigma<- sqrt(25) 
u <- 22
```

i. lies between 16.2 and 27.5

```{r}

pnorm(27.5, mean=u, sd=sigma) - pnorm(16.2, mean=u, sd=sigma)

```

ii. is greater than 29

```{r}

1-pnorm(29, mean=u, sd=sigma)

```

iii. is less than 17

```{r}

pnorm(17, mean=u, sd=sigma)

```

iv. is less than 15 or greater than 25

```{r}

(1-pnorm(25, mean=u, sd=sigma)) + pnorm(15, mean=u, sd=sigma)

```

#### b) Probability that in 60 tosses of a fair coin the head comes up

$P(X=k) = {n\choose k} \cdot p^kq^{n-k}$

$p=0.5$ and $q=0.5$

i. 20, 25 or 30 times

```{r}

dbinom(20, size=60, prob=0.5) + dbinom(25, size=60, prob=0.5) + dbinom(30, size=60, prob=0.5)

```

ii. less than 20 times

```{r}

pbinom(20, size=60, prob=0.5)

```

iii. between 20 and 30 times

```{r}

pbinom(30, size=60, prob=0.5) - pbinom(20, size=60, prob=0.5)

```

#### c) A random variable X has Poisson distribution with mean 7. Find the probability that

$P\left( x \right) = \frac{{e^{ - \lambda } \lambda ^x }}{{x!}}$

i. X is less than 5

```{r}

ppois(5, lambda=7)

```

ii. X is greater than 10 (strictly)

```{r}

1 - ppois(10, lambda=7)

```

iii. X is between 4 and 16

```{r}

ppois(16, lambda=7) - ppois(4, lambda=7)

```
